\phantomsection\addcontentsline{toc}{section}{Digital Twin}
\ECchapter{21}{Digital Twin}{Creating virtual replicas of physical systems}{ECBlue}

\ECObjectives{
\begin{itemize}
\item Explain how measurements update a digital model.
\item Distinguish simulation, monitoring and prediction.
\item Identify the limits of a digital twin.
\end{itemize}}

\begin{multicols}{2}
\begin{engineeringnews}
Digital twins are increasingly used for aircraft engines, factories, bridges and energy systems. By
combining sensor data with physical or statistical models, engineers can estimate internal states,
predict degradation and test operating strategies without disturbing the real asset.
\end{engineeringnews}

\begin{ecarticle}
A digital twin is a virtual representation of a physical system that is regularly updated with real
measurements. Unlike a static CAD model, it evolves with the asset. Sensors measure quantities such as
temperature, vibration, force, pressure or electrical current. These data are filtered and compared
with a mathematical model that describes how the system should behave.

The model may be physics-based, data-driven or hybrid. A physics-based model uses conservation laws,
material properties and geometry. A data-driven model learns relationships from historical data. A
hybrid twin combines both approaches: physics constrains the solution while machine learning corrects
unknown effects or accelerates computation.

Engineers use twins for three main purposes. First, monitoring compares measured and expected
behaviour. A growing difference can reveal sensor drift, wear or an incorrect model. Second,
prediction estimates future states, for example remaining bearing life or battery temperature during
a demanding mission. Third, optimisation allows engineers to test control settings, maintenance dates
or production schedules before applying them to the real system.

A useful twin depends on data quality. Measurements must be time-stamped, calibrated and transmitted
reliably. The model must also be validated over the full operating range. A twin that performs well at
nominal speed may fail during start-up, overload or extreme weather. Uncertainty must therefore be
reported rather than hidden.

Digital twins do not replace physical tests. They help engineers decide which tests are most useful
and interpret measurements more effectively. Their value comes from the continuous loop between the
real asset, the data platform and the model.
\end{ecarticle}

\begin{tcolorbox}[title=\sffamily\bfseries Did You Know?,colback=yellow!10,colframe=ECOrange]
A model becomes a true digital twin only when information flows repeatedly between the physical asset
and its virtual representation. Without this update loop, it is simply a simulation model.
\end{tcolorbox}

\begin{engineeringfocus}
\textbf{Information loop}
\begin{center}
\begin{tikzpicture}[font=\scriptsize,>=Latex,node distance=4mm]
\node[draw=ECBlue,fill=ECLightBlue,rounded corners,align=center] (asset) {physical\\asset};
\node[draw=ECGreen,fill=ECGreen!10,rounded corners,align=center,right=of asset] (data) {sensors and\\data platform};
\node[draw=ECPurple,fill=ECPurple!10,rounded corners,align=center,right=of data] (twin) {digital\\twin};
\node[draw=ECOrange,fill=ECOrange!10,rounded corners,align=center,below=of data] (decision) {diagnosis and\\decision};
\draw[->,thick] (asset)-- node[above]{measurements} (data);
\draw[->,thick] (data)-- node[above]{updates} (twin);
\draw[->,thick] (twin)--(decision);
\draw[->,thick] (decision)--(asset);
\end{tikzpicture}
\end{center}
\end{engineeringfocus}

\begin{keyfigures}
\begin{tabularx}{\linewidth}{@{}Xr@{}}
Essential input & calibrated data\\
Core operation & model update\\
Typical output & health estimate\\
Main risk & model drift\\
Validation method & test comparison
\end{tabularx}
\end{keyfigures}

\begin{tcolorbox}[title=\sffamily\bfseries Prediction error over time,colback=white,colframe=ECBlue]
\begin{center}
\begin{tikzpicture}
\begin{axis}[width=\linewidth,height=48mm,xlabel={Time (days)},ylabel={Error (\%)},grid=major,
ticklabel style={font=\scriptsize}]
\addplot[thick,mark=*] table[x=time,y=error,col sep=comma]{data/EC21/prediction.csv};
\end{axis}
\end{tikzpicture}
\end{center}
\footnotesize A rising residual may indicate wear, a changing environment or an outdated model.
\end{tcolorbox}

\begin{vocabulary}
\begin{tabularx}{\linewidth}{@{}lX@{}}
digital twin & jumeau numérique\\
asset & équipement réel\\
state estimation & estimation d'état\\
residual & résidu / écart\\
calibration & étalonnage\\
model drift & dérive du modèle\\
remaining useful life & durée de vie restante\\
uncertainty & incertitude\\
data-driven & fondé sur les données\\
validation & validation
\end{tabularx}
\end{vocabulary}

\begin{questions}
\item How is a digital twin different from a CAD model?
\item What is the role of sensors in the update loop?
\item Compare physics-based and data-driven models.
\item Why must a twin be validated outside nominal operation?
\item What can cause the prediction error to increase?
\end{questions}

\begin{engineeringchallenge}
Design a digital twin for a school ventilation unit. Select four sensors, define the model outputs,
propose one fault indicator and explain how the twin could reduce energy use without lowering air
quality.
\end{engineeringchallenge}

\begin{speaking}
Should safety-critical decisions ever be made automatically by a digital twin? Discuss reliability,
responsibility, human supervision and uncertainty.
\end{speaking}
\end{multicols}

\subsection*{Sources}
\ECsource{NIST -- Digital Twins}{https://www.nist.gov/topics/digital-twins}
\ECsource{NASA -- Digital Twin}{https://ntrs.nasa.gov/}
