\phantomsection\addcontentsline{toc}{section}{Autonomous Vehicles}
\ECchapter{23}{Autonomous Vehicles}{How can an autonomous vehicle remain safe?}{ECOrange}

\ECObjectives{
\begin{itemize}
\item Describe perception, localisation, prediction, planning and control.
\item Explain sensor diversity, redundancy and degraded operating modes.
\item Calculate stopping distance and define an operational design domain.
\end{itemize}}

\begin{multicols}{2}
\begin{engineeringnews}
Autonomous systems are being tested in passenger cars, shuttles, mines, ports and warehouses. Progress
is fastest in controlled environments, while public roads remain difficult because weather, road
geometry, human behaviour and rare events vary widely.
\end{engineeringnews}

\begin{ecarticle}
An autonomous vehicle must perceive its environment, estimate its own position, predict how other road
users may move, plan a safe trajectory and control steering, braking and propulsion. These functions run
continuously under strict timing, reliability and safety constraints.

Perception combines cameras, radar, lidar and ultrasonic sensors. Cameras provide colour and detailed
shapes but can be affected by darkness, glare or fog. Radar measures distance and relative speed. Lidar
produces accurate three-dimensional point clouds but may be sensitive to rain or contamination. Sensor
fusion combines complementary information and helps detect inconsistent measurements.

Localisation uses satellite positioning, inertial sensors, wheel odometry and maps. No source is
reliable everywhere: GNSS can be blocked in tunnels, while wheel slip creates odometry errors. Redundant
estimates allow the system to identify faulty data and enter a safe degraded mode.

The prediction module estimates the future motion of pedestrians, cyclists and vehicles. The planner
selects a manoeuvre and generates a trajectory that respects road geometry, traffic rules, vehicle
dynamics and comfort. The controller then adjusts steering torque, brake pressure and propulsion.

Total stopping distance includes the distance travelled during sensing, computation and actuator delay,
plus the physical braking distance. Because kinetic energy is proportional to $v^2$, braking distance
rises rapidly with speed. Wet roads, worn tyres and downhill gradients reduce available deceleration.

Automation must operate inside a defined operational design domain specifying road type, speed, weather,
visibility and map quality. At its boundary, the vehicle must request human control or perform a
minimal-risk manoeuvre.
\end{ecarticle}

\begin{tcolorbox}[title=\sffamily\bfseries Engineering Insight,colback=yellow!10,colframe=ECOrange]
Correct object detection is not enough. Safety must be verified across the complete chain from sensing
to prediction, decision, control and fallback behaviour.
\end{tcolorbox}

\begin{engineeringfocus}
\textbf{Autonomous driving pipeline}
\begin{center}
\resizebox{.98\linewidth}{!}{%
\begin{tikzpicture}[font=\scriptsize,>=Latex,node distance=3mm]
\node[draw=ECBlue,fill=ECLightBlue,rounded corners] (sense) {sensors};
\node[draw=ECGreen,fill=ECGreen!10,rounded corners,right=of sense] (per) {perception};
\node[draw=ECPurple,fill=ECPurple!10,rounded corners,right=of per] (pred) {prediction};
\node[draw=ECOrange,fill=ECOrange!10,rounded corners,right=of pred] (plan) {planning};
\node[draw=ECRed,fill=ECRed!8,rounded corners,right=of plan] (ctrl) {control};
\node[draw=black,fill=gray!10,rounded corners,below=of plan] (safe) {safety monitor};
\draw[->,thick] (sense)--(per); \draw[->,thick] (per)--(pred); \draw[->,thick] (pred)--(plan); \draw[->,thick] (plan)--(ctrl);
\draw[->,thick,dashed] (safe)--(ctrl); \draw[->,thick,dashed] (per)--(safe);
\end{tikzpicture}%
}
\end{center}
The safety monitor checks plausibility, timing and operational-domain limits independently of the nominal planner.
\end{engineeringfocus}

\begin{keyfigures}
\begin{tabularx}{\linewidth}{@{}Xr@{}}
Camera strength & detailed images\\
Radar strength & relative speed\\
Lidar output & 3D point cloud\\
Critical principle & redundancy\\
Safe fallback & minimal-risk manoeuvre
\end{tabularx}
\end{keyfigures}

\begin{tcolorbox}[title=\sffamily\bfseries Illustrative braking distance,colback=white,colframe=ECOrange]
\begin{center}
\begin{tikzpicture}
\begin{axis}[width=.96\linewidth,height=46mm,xlabel={Speed (km/h)},ylabel={Braking distance (m)},grid=major,legend style={font=\scriptsize,at={(.5,-.22)},anchor=north,legend columns=2},ticklabel style={font=\scriptsize}]
\addplot[thick,mark=*] table[x=speed,y=dry,col sep=comma]{data/EC23/stopping.csv};
\addplot[thick,mark=square*] table[x=speed,y=wet,col sep=comma]{data/EC23/stopping.csv};
\legend{Dry road,Wet road}
\end{axis}
\end{tikzpicture}
\end{center}
\end{tcolorbox}

\begin{vocabulary}
\begin{tabularx}{\linewidth}{@{}lX@{}}
sensor fusion & fusion de capteurs\\
point cloud & nuage de points\\
localisation & localisation\\
trajectory & trajectoire\\
fallback & solution de repli\\
operational design domain & domaine opérationnel\\
wheel slip & glissement des roues\\
braking distance & distance de freinage\\
redundancy & redondance\\
minimal-risk manoeuvre & manœuvre de risque minimal
\end{tabularx}
\end{vocabulary}

\begin{questions}
\item What functions form the autonomous-driving chain?
\item Why are several sensor technologies combined?
\item How can localisation fail in a tunnel?
\item Why does braking distance rise rapidly with speed?
\item What information belongs in an operational design domain?
\end{questions}

\begin{applicationexercise}
A shuttle travels at $36\,\mathrm{km\,h^{-1}}$. Sensing and computation take $0.35\,\mathrm{s}$ and
the brake actuator adds $0.15\,\mathrm{s}$. Assume a deceleration of $5.0\,\mathrm{m\,s^{-2}}$.
\begin{enumerate}
\item Convert the speed into $\mathrm{m\,s^{-1}}$.
\item Calculate the distance travelled during the total delay.
\item Using $d_b=v^2/(2a)$, calculate the braking distance.
\item Determine the total stopping distance and a detection range with a $25\%$ safety margin.
\item Repeat the result for a degraded deceleration of $3.0\,\mathrm{m\,s^{-2}}$.
\end{enumerate}
\end{applicationexercise}

\begin{engineeringchallenge}
Design an autonomous shuttle for a closed university campus. Define its sensors, maximum speed,
operational design domain and minimal-risk manoeuvre. Add a verification matrix covering pedestrians,
rain, glare, GNSS loss and one faulty sensor, with a measurable acceptance criterion for each test.
\end{engineeringchallenge}

\begin{speaking}
Who should be responsible after an autonomous-vehicle accident: the passenger, manufacturer, software
provider or road operator? Defend a position using technical evidence.
\end{speaking}
\end{multicols}

\subsection*{Sources}
\ECsource{SAE International -- Automated Driving}{https://www.sae.org/standards/content/j3016_202104/}
\ECsource{NHTSA -- Automated Vehicles}{https://www.nhtsa.gov/vehicle-safety/automated-vehicles-safety}