\phantomsection\addcontentsline{toc}{section}{Smart Grids}
\ECchapter{24}{Smart Grids}{Managing tomorrow's electricity networks}{ECPurple}

\ECObjectives{
\begin{itemize}
\item Explain how generation and demand must remain balanced.
\item Describe the role of storage, forecasting and demand response.
\item Identify resilience and cybersecurity constraints.
\end{itemize}}

\begin{multicols}{2}
\begin{engineeringnews}
Electricity networks are changing from one-way systems dominated by large power stations into
interactive networks containing solar roofs, wind farms, batteries, electric vehicles and flexible
loads. Digital control is becoming essential to coordinate these distributed resources.
\end{engineeringnews}

\begin{ecarticle}
An electricity grid must balance production and consumption at every moment. If generation is lower
than demand, frequency falls; if it is higher, frequency rises. Traditional grids adjusted a limited
number of controllable power stations. Smart grids coordinate many smaller and more variable devices.

Smart meters measure energy flows at short intervals. Forecasting software estimates demand and
renewable generation. Automated substations can change network configuration after a fault, while
batteries and flexible loads respond rapidly to imbalances. Demand response shifts selected uses in
time: for example, water heating or electric-vehicle charging may be delayed when the grid is heavily
loaded.

Distributed generation creates bidirectional power flows. A neighbourhood may consume electricity in
the evening but export solar power at midday. Voltage can rise near photovoltaic installations, and
protection systems must distinguish normal reverse flow from a dangerous fault. Engineers therefore
combine power electronics, communication systems and network models.

Storage provides several services. Batteries can smooth short fluctuations, support frequency and
reduce peak demand. Pumped hydro stores larger amounts of energy for longer periods. Electric vehicles
may eventually provide controlled charging or energy back to the grid, but battery degradation and
user mobility must be respected.

Digitalisation increases exposure to cyberattack and communication failure. Local controllers must
retain safe behaviour when the central platform is unavailable. Resilience also requires physical
redundancy, black-start capability and plans for extreme weather. A smart grid is not merely connected;
it must remain stable, secure and understandable during abnormal conditions.
\end{ecarticle}

\begin{tcolorbox}[title=\sffamily\bfseries Did You Know?,colback=yellow!10,colframe=ECOrange]
Electricity itself travels through the network almost instantly, but energy planning may cover seconds,
hours, seasons and years. Different storage technologies are needed for these different timescales.
\end{tcolorbox}

\begin{engineeringfocus}
\textbf{Bidirectional smart grid}
\begin{center}
\begin{tikzpicture}[font=\scriptsize,>=Latex,node distance=4mm]
\node[draw=ECPurple,fill=ECPurple!10,rounded corners] (grid) {distribution grid};
\node[draw=ECOrange,fill=ECOrange!10,rounded corners,left=of grid] (solar) {solar and wind};
\node[draw=ECGreen,fill=ECGreen!10,rounded corners,right=of grid] (loads) {homes and industry};
\node[draw=ECBlue,fill=ECLightBlue,rounded corners,below=of grid] (store) {storage and EVs};
\node[draw=ECRed,fill=ECRed!8,rounded corners,above=of grid] (control) {forecast and control};
\draw[<->,thick] (solar)--(grid); \draw[<->,thick] (grid)--(loads); \draw[<->,thick] (grid)--(store);
\draw[<->,thick,dashed] (control)--(grid);
\end{tikzpicture}
\end{center}
\end{engineeringfocus}

\begin{keyfigures}
\begin{tabularx}{\linewidth}{@{}Xr@{}}
Instant requirement & power balance\\
Fast flexibility & batteries\\
Long storage & pumped hydro\\
Flexible demand & smart charging\\
Critical risk & cyberattack
\end{tabularx}
\end{keyfigures}

\begin{tcolorbox}[title=\sffamily\bfseries Daily balancing example,colback=white,colframe=ECPurple]
\begin{center}
\begin{tikzpicture}
\begin{axis}[width=\linewidth,height=50mm,xlabel={Hour},ylabel={Power (MW)},grid=major,
legend style={font=\scriptsize,at={(0.5,-0.2)},anchor=north,legend columns=3},ticklabel style={font=\scriptsize}]
\addplot[thick] table[x=hour,y=demand,col sep=comma]{data/EC24/load.csv};
\addplot[thick] table[x=hour,y=solar,col sep=comma]{data/EC24/load.csv};
\addplot[thick,dashed] table[x=hour,y=battery,col sep=comma]{data/EC24/load.csv};
\legend{Demand,Solar,Battery output}
\end{axis}
\end{tikzpicture}
\end{center}
\end{tcolorbox}

\begin{vocabulary}
\begin{tabularx}{\linewidth}{@{}lX@{}}
smart meter & compteur communicant\\
demand response & effacement / flexibilité\\
distributed generation & production décentralisée\\
peak demand & pointe de consommation\\
frequency control & réglage de fréquence\\
bidirectional flow & flux bidirectionnel\\
substation & poste électrique\\
black start & redémarrage autonome\\
load shedding & délestage\\
grid resilience & résilience du réseau
\end{tabularx}
\end{vocabulary}

\begin{questions}
\item Why must generation equal demand continuously?
\item How does demand response help the network?
\item What problem can midday solar export create?
\item Compare batteries and pumped hydro.
\item Why should local controllers work without the central platform?
\end{questions}

\begin{engineeringchallenge}
Plan the energy management of a school with rooftop solar panels, a 200 kWh battery and ten EV
chargers. Define priorities for a normal day, a winter peak and a two-hour grid outage.
\end{engineeringchallenge}

\begin{speaking}
Should electricity suppliers be allowed to delay domestic appliances automatically during peak demand?
Discuss consent, comfort, price, privacy and grid stability.
\end{speaking}
\end{multicols}

\subsection*{Sources}
\ECsource{International Energy Agency -- Smart Grids}{https://www.iea.org/energy-system/electricity/smart-grids}
\ECsource{ENTSO-E}{https://www.entsoe.eu/}
