\phantomsection\addcontentsline{toc}{section}{Biomedical Imaging}
\ECchapter{26}{Biomedical Imaging}{Seeing inside the human body}{ECRed}

\ECObjectives{
\begin{itemize}
\item Explain the physical principles of the main imaging techniques.
\item Interpret resolution, contrast, noise and acquisition trade-offs.
\item Estimate simple imaging quantities and identify safety constraints.
\end{itemize}}

\begin{multicols}{2}
\begin{engineeringnews}
Biomedical imaging allows clinicians to observe anatomy and function without surgery. Engineers combine
wave physics, detectors, electronics, signal processing and reconstruction algorithms to produce clinically
useful images while controlling acquisition time, cost and patient risk.
\end{engineeringnews}

\begin{ecarticle}
Biomedical imaging converts physical interactions inside the body into maps that clinicians can interpret.
Different technologies measure different properties. X-ray radiography and computed tomography use the
attenuation of ionising radiation. Ultrasound sends mechanical waves and analyses their echoes. Magnetic
resonance imaging, or MRI, measures signals produced by hydrogen nuclei in a strong magnetic field. Nuclear
medicine detects radiation emitted by a tracer introduced into the body.

Every imaging chain contains four main stages: excitation, interaction, detection and reconstruction. The
detector rarely produces a ready-to-use picture. It records many noisy signals from different positions or
times. A reconstruction algorithm estimates the spatial distribution that best explains these measurements.
Increasing spatial resolution generally requires more samples, a stronger signal or a longer acquisition.

Image quality is described by spatial resolution, contrast, signal-to-noise ratio and artefacts. Resolution
determines the smallest visible detail. Contrast separates tissues with similar properties. Noise creates
random variations, whereas artefacts create structured errors caused by motion, metal objects, calibration
errors or incomplete measurements. Engineers must select acquisition parameters that balance these effects.

In ultrasound, depth can be estimated from the echo round-trip time:
\[
d=\frac{ct}{2},
\]
where $c$ is the speed of sound in tissue and $t$ is the measured delay. The factor two appears because the
pulse travels to the reflector and back to the probe. Shorter pulses improve axial resolution, but reducing
the transmitted energy may lower the signal-to-noise ratio.

Safety is equally important. X-ray dose must be kept as low as reasonably achievable. MRI avoids ionising
radiation but requires strict control of magnetic objects, radio-frequency heating and acoustic noise.
Ultrasound is portable and relatively inexpensive, yet image quality depends strongly on the operator and
on acoustic access to the organ.

Artificial intelligence can assist segmentation, reconstruction and anomaly detection, but it must be
validated on diverse patients and scanners. A convincing image is not necessarily an accurate one.
Biomedical engineers therefore work closely with physicists, radiologists, clinicians and regulators.
\end{ecarticle}

\begin{tcolorbox}[title=\sffamily\bfseries Did You Know?,colback=yellow!10,colframe=ECOrange]
A high-performance prototype is not sufficient: engineers must demonstrate repeatability, safety and useful
performance under realistic clinical conditions.
\end{tcolorbox}

\begin{engineeringfocus}
\textbf{From measurement to diagnosis}
\begin{center}
\resizebox{0.96\linewidth}{!}{%
\begin{tikzpicture}[font=\scriptsize,>=Latex,node distance=4mm]
\node[draw=ECRed,fill=ECRed!10,rounded corners] (source) {energy source};
\node[draw=ECOrange,fill=ECOrange!10,rounded corners,right=of source] (body) {body interaction};
\node[draw=ECBlue,fill=ECLightBlue,rounded corners,right=of body] (det) {detector};
\node[draw=ECPurple,fill=ECPurple!10,rounded corners,below=of body] (alg) {reconstruction};
\node[draw=ECGreen,fill=ECGreen!10,rounded corners,below=of det] (image) {clinical image};
\draw[->,thick] (source)--(body); \draw[->,thick] (body)--(det); \draw[->,thick] (det)--(alg); \draw[->,thick] (alg)--(image);
\end{tikzpicture}}
\end{center}
\end{engineeringfocus}

\begin{keyfigures}
\begin{tabularx}{\linewidth}{@{}Xr@{}}
X-ray / CT & attenuation\\
Ultrasound & echo delay\\
MRI & magnetic resonance\\
Main trade-off & quality vs risk\\
Critical step & validation
\end{tabularx}
\end{keyfigures}

\begin{tcolorbox}[title=\sffamily\bfseries Resolution and acquisition time,colback=white,colframe=ECRed]
\begin{center}
\begin{tikzpicture}
\begin{axis}[width=0.96\linewidth,height=46mm,xlabel={Voxel size (mm)},ylabel={Acquisition time (s)},
grid=major,ticklabel style={font=\scriptsize},x dir=reverse]
\addplot[thick,mark=*] table[x=resolution,y=time,col sep=comma]{data/EC26/imaging.csv};
\end{axis}
\end{tikzpicture}
\end{center}
\end{tcolorbox}

\begin{tcolorbox}[title=\sffamily\bfseries Quantitative application,colback=ECRed!4,colframe=ECRed]
An ultrasound pulse returns from a tissue boundary after $130\,\mu\mathrm{s}$. Take the speed of sound in
soft tissue as $1540\,\mathrm{m\,s^{-1}}$.
\begin{enumerate}
\item Calculate the depth of the boundary.
\item A second echo arrives $26\,\mu\mathrm{s}$ later. Estimate the distance between the two boundaries.
\item Explain why using the full travel distance instead of half would give an incorrect result.
\end{enumerate}
\end{tcolorbox}

\begin{vocabulary}
\begin{tabularx}{\linewidth}{@{}lX@{}}
attenuation & atténuation\\
echo & écho\\
voxel & voxel\\
spatial resolution & résolution spatiale\\
contrast & contraste\\
noise & bruit\\
artefact & artefact\\
reconstruction & reconstruction\\
radiation dose & dose de rayonnement\\
tracer & traceur
\end{tabularx}
\end{vocabulary}

\begin{questions}
\item Why do different imaging technologies reveal different tissues?
\item What is the role of reconstruction algorithms?
\item Explain the difference between noise and an artefact.
\item Why is higher spatial resolution not always the best choice?
\item Give one safety constraint for CT and one for MRI.
\end{questions}

\begin{engineeringchallenge}
Design an imaging protocol for a moving organ. Choose a technology, define the required resolution and
acquisition time, identify one major artefact and propose a method to reduce it.
\end{engineeringchallenge}

\begin{speaking}
Should AI-generated medical images be accepted when they improve visibility but modify the original
measurements? Discuss accuracy, responsibility and patient safety.
\end{speaking}
\end{multicols}

\subsection*{Sources}
\ECsource{World Health Organization -- Medical imaging}{https://www.who.int/health-topics/medical-devices}
\ECsource{International Atomic Energy Agency -- Radiation protection}{https://www.iaea.org/resources/rpop}