\phantomsection\addcontentsline{toc}{section}{Quantum Technologies}
\ECchapter{28}{Quantum Technologies}{Controlling fragile states for computing, sensing and communication}{ECPurple}

\ECObjectives{
\begin{itemize}
\item Distinguish classical bits from qubits and explain superposition, interference and measurement.
\item Analyse coherence time, gate duration, error rate and hardware overhead.
\item Compare realistic applications in quantum computing, sensing and secure communication.
\end{itemize}}

\begin{multicols}{2}
\begin{engineeringnews}
Quantum technologies are progressing from laboratory demonstrations towards specialised sensors, secure links
and prototype processors. Their engineering value comes from preparing, controlling and measuring individual
quantum states with sufficient fidelity, stability and repeatability.
\end{engineeringnews}

\begin{ecarticle}
A classical bit takes the value 0 or 1. A qubit can be prepared in a state
$\alpha\lvert0\rangle+\beta\lvert1\rangle$, where $|\alpha|^2+|\beta|^2=1$. Measurement returns 0 or 1 with
probabilities determined by these amplitudes. The amplitudes themselves are not directly observable in a single
measurement, so a quantum experiment must be repeated many times to estimate probabilities.

Quantum computation does not simply test every answer at once. Gates modify amplitudes, while interference
suppresses some outcomes and reinforces others. A useful algorithm must therefore organise interference so that
informative results become more probable before the final measurement. Quantum speed-up is task-dependent and
does not make every computation faster.

Several qubits may become entangled, meaning that their joint state cannot be represented as independent classical
variables. Entanglement is a resource for algorithms, communication and metrology, but it also increases the
complexity of calibration and verification.

The main engineering limitation is decoherence. Unwanted interactions with the environment disturb phase and
energy states. Superconducting processors may require millikelvin temperatures and filtered microwave control.
Trapped-ion and neutral-atom systems require ultra-high vacuum, stable lasers and precise electromagnetic fields.
Photonic systems avoid some cooling constraints but need low-loss sources, circuits and detectors.

If coherence decays exponentially, a simple model is
\[
C(t)=C_0\exp(-t/T_2),
\]
where $T_2$ is the coherence time. This model gives a first estimate of the available operation window, but a real
processor is also limited by gate errors, readout errors, cross-talk, leakage and calibration drift.

Quantum error correction encodes one logical qubit in many physical qubits. Syndrome measurements reveal selected
errors without directly measuring the logical information. Fault-tolerant operation requires physical error rates
below a demanding threshold and a large hardware overhead. Qubit count alone is therefore a poor performance metric;
fidelity, connectivity, cycle time and logical error rate are equally important.

Quantum sensing may reach practical use earlier in several fields. Atomic clocks, magnetometers and gravimeters
measure tiny frequency, magnetic-field or acceleration changes. Longer interrogation can improve sensitivity, but
also increases exposure to decoherence and environmental drift. Engineering optimisation must include duty cycle,
calibration and conventional signal processing.
\end{ecarticle}

\begin{tcolorbox}[title=\sffamily\bfseries Engineering insight,colback=yellow!10,colframe=ECOrange]
A large number of physical qubits does not guarantee useful computation. The relevant question is how many reliable
logical operations can be completed before accumulated error makes the result unusable.
\end{tcolorbox}

\begin{engineeringfocus}
\textbf{Hybrid quantum--classical control loop}
\begin{center}
\resizebox{0.96\linewidth}{!}{%
\begin{tikzpicture}[font=\scriptsize,>=Latex,node distance=3.5mm]
\node[draw=ECPurple,fill=ECPurple!10,rounded corners] (init) {prepare state};
\node[draw=ECBlue,fill=ECLightBlue,rounded corners,right=of init] (gates) {apply gates};
\node[draw=ECOrange,fill=ECOrange!10,rounded corners,right=of gates] (meas) {measure};
\node[draw=ECGreen,fill=ECGreen!10,rounded corners,below=of gates] (class) {classical estimation};
\node[draw=ECRed,fill=ECRed!8,rounded corners,below=of meas] (cal) {calibration monitor};
\draw[->,thick] (init)--(gates); \draw[->,thick] (gates)--(meas);
\draw[->,thick] (meas)--(class); \draw[->,thick] (class)--(gates);
\draw[->,thick,dashed] (cal)--(gates); \draw[->,thick,dashed] (meas)--(cal);
\end{tikzpicture}}
\end{center}
\end{engineeringfocus}

\begin{keyfigures}
\begin{tabularx}{\linewidth}{@{}Xr@{}}
Information unit & qubit\\
State lifetime & coherence time\\
Operation quality & gate fidelity\\
Protected information & logical qubit\\
Near-term strength & sensing
\end{tabularx}
\end{keyfigures}

\begin{tcolorbox}[title=\sffamily\bfseries Illustrative coherence decay,colback=white,colframe=ECPurple]
\begin{center}
\begin{tikzpicture}
\begin{axis}[width=0.94\linewidth,height=45mm,xlabel={Time (arbitrary units)},ylabel={Remaining coherence},
ymin=0,ymax=1.05,grid=major,ticklabel style={font=\scriptsize},label style={font=\scriptsize}]
\addplot[thick,mark=*] table[x=time,y=coherence,col sep=comma]{data/EC28/coherence.csv};
\end{axis}
\end{tikzpicture}
\end{center}
\footnotesize The exponential model is useful for first estimates but does not represent every noise mechanism.
\end{tcolorbox}

\begin{applicationexercise}{Coherence budget}
A qubit has $C_0=1$ and $T_2=80\,\mu\mathrm{s}$. Assume $C(t)=\exp(-t/T_2)$.
\begin{enumerate}
\item Calculate the remaining coherence after $40\,\mu\mathrm{s}$ and $120\,\mu\mathrm{s}$.
\item Determine the maximum time for which $C(t)\geq0.70$.
\item A gate lasts $50\,\mathrm{ns}$. Estimate the theoretical number of sequential gates in this interval.
\item If each gate has fidelity 99.9\%, estimate the probability that 500 gates are all error-free.
\item Explain why both answers are optimistic upper bounds.
\end{enumerate}
\end{applicationexercise}

\begin{vocabulary}
\begin{tabularx}{\linewidth}{@{}lX@{}}
qubit & qubit\\
superposition & superposition\\
entanglement & intrication\\
interference & interférence\\
measurement & mesure\\
decoherence & décohérence\\
quantum gate & porte quantique\\
error correction & correction d’erreurs\\
gate fidelity & fidélité d’une porte\\
fault-tolerant & tolérant aux fautes
\end{tabularx}
\end{vocabulary}

\begin{questions}
\item How does a qubit differ from a classical bit?
\item Why must a quantum experiment be repeated to estimate probabilities?
\item What physical mechanisms can cause decoherence?
\item Why are many physical qubits needed for one logical qubit?
\item Why is qubit count alone an insufficient benchmark?
\end{questions}

\begin{engineeringchallenge}
Compare three candidate technologies for a quantum sensor: superconducting circuits, trapped atoms and photonics.
Define five weighted criteria, including sensitivity, operating environment, calibration, maturity and cost. Build a
decision matrix, identify one dominant risk for the selected technology and propose a validation experiment with an
acceptance threshold and a conventional reference sensor.
\end{engineeringchallenge}

\begin{speaking}
Should governments prioritise quantum computing, quantum sensing or quantum-secure communication? Defend one
investment strategy using maturity, expected social benefit, infrastructure cost and technological sovereignty.
\end{speaking}
\end{multicols}

\subsection*{Sources}
\ECsource{NIST -- Quantum information science}{https://www.nist.gov/topics/quantum-information-science}
\ECsource{European Quantum Flagship}{https://qt.eu/}